---
title: "aad-filelist"
subtitle: "Antarctic / Southern Ocean file-listing cache"
format:
  html:
    theme: cosmo
    toc: false
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(purrr)
library(arrow)
library(glue)
library(fs)
library(bowerbird)
## GitHub repo details -------------------------------------------------------
gh_owner <- "mdsumner"
gh_repo  <- "aad-filelist"
gh_raw   <- glue("https://raw.githubusercontent.com/{gh_owner}/{gh_repo}/main")
## Release assets are at:
## https://github.com/{owner}/{repo}/releases/download/latest/{id}.parquet
release_base <- glue("https://github.com/{gh_owner}/{gh_repo}/releases/download/latest")
```

```{r load-sources}
## Source each sources/*.R to extract metadata without running a sync.
## Each file defines a single `src` object (a bb_source tibble).
source_files <- fs::dir_ls("sources", glob = "*.R")

## load-sources chunk — add data_group to the extract
source_meta <- map_dfr(source_files, function(f) {
  e <- new.env(parent = baseenv())
  suppressPackageStartupMessages(source(f, local = e))
  s <- e$src
  tibble(
    source_file  = fs::path_file(f),
    id           = s$id[[1]],
    name         = s$name[[1]],
    description  = s$description[[1]],
    doc_url      = s$doc_url[[1]],
    license      = s$license[[1]],
    data_group   = if ("data_group" %in% names(s)) s$data_group[[1]] else "Ungrouped"
  )
})
```

```{r read-parquets, eval=TRUE}
#files <- list.files("sources", pattern = "\\.parquet$", full.names = TRUE)  # your actual glob
#cat("found", length(files), "parquet(s):\n"); print(files)
#stopifnot("no parquet files on runner" = length(files) > 0)

## Try to read the first row of each Parquet asset to get last sync time.
## Falls back gracefully if the asset doesn't exist yet.
get_parquet_meta <- function(id) {
  url <- glue("{release_base}/{id}.parquet")
  result <- tryCatch({
    df <- read_parquet(url)
    list(
      n_files     = nrow(df),
      sync_time   = format(max(df$sync_time, na.rm = TRUE), "%Y-%m-%d %H:%M UTC"),
      parquet_url = url
    )
  }, error = function(e) {
    list(
      n_files     = NA_integer_,
      sync_time   = "not yet synced",
      parquet_url = url
    )
  })
  result
}

parquet_meta <- map(source_meta$id, get_parquet_meta)

## extract each field as a plain vector rather than relying on map_int/map_chr
## on a list-of-lists, which is fragile when fallbacks are involved
summary_tbl <- source_meta |>
  mutate(
    n_files     = vapply(parquet_meta, `[[`, integer(1), "n_files"),
    sync_time   = vapply(parquet_meta, `[[`, character(1), "sync_time"),
    parquet_url = vapply(parquet_meta, `[[`, character(1), "parquet_url"),
    source_url  = glue("{gh_raw}/sources/{source_file}")
  )
```

This repo provides per-collection file-listing caches for Antarctic regional
climate model output.  Each collection is defined in its own R file using
[bowerbird](https://docs.ropensci.org/bowerbird/), and the resulting file
manifest is published here as a Parquet file for each release.


## Individual manifests



**To use a manifest in R:**

```r
arrow::read_parquet(
  "https://github.com/mdsumner/aad-filelist/releases/download/latest/cordex-polarres-ant12-hclim-dmi.parquet"
)
```

---

```{r table, eval=TRUE, results="asis"}
## table chunk
groups <- split(summary_tbl, summary_tbl$data_group)

for (grp_name in sort(names(groups))) {
  cat("##", grp_name, "\n\n")
  
  groups[[grp_name]] |>
    transmute(
      Collection     = glue('<a href="{doc_url}">{name}</a>'),
      Description    = description,
      `R source`     = glue('<a href="{source_url}"><code>{source_file}</code></a>'),
      `Parquet cache`= glue('<a href="{parquet_url}"><code>{id}.parquet</code></a>'),
      `Files listed` = ifelse(is.na(n_files), "—", as.character(n_files)),
      `Last synced`  = sync_time
    ) |>
    knitr::kable(format = "html", escape = FALSE) |>
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
                              full_width = TRUE) |>
    print()
  
  cat("\n\n")
}
```


```{r read-full-raad,eval=TRUE}
raad_url <- "https://github.com/mdsumner/aad-filelist/releases/download/latest/raad_file_db.parquet"
raad_df <- NULL
raad_df <- tryCatch(arrow::read_parquet(raad_url), error = function(e) NULL)

if (!is.null(raad_df)) {
  ## extract top-level domain / data provider from the file path
  raad_summary <- raad_df |>
    mutate(provider = stringr::str_extract(file, "^[^/]+")) |>
    count(provider, name = "n_files") |>
    filter(n_files > 1) |> 
    arrange(desc(n_files))
}
```
## Full bowerbird+raadtools PUBLIC list

```{r raad-table, results="asis"}
if (!is.null(raad_df)) {
 # cat(glue("**{nrow(raad_df):,} files** across {nrow(raad_summary)} providers,",
#           " last updated from the RAAD server.\n\n"))

  raad_summary |>
    mutate(
      `Data provider` = glue('<a href="https://{provider}">{provider}</a>'),
      `Files in RAAD` = scales::comma(n_files)
    ) |>
    select(`Data provider`, `Files in RAAD`) |>
    knitr::kable(format = "html", escape = FALSE) |>
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
                              full_width = FALSE) |>
    print()
}
```

Full listing as Parquet (1.4M rows, `file` column, prepend `https://` for URL):

```r
raad_df <- arrow::read_parquet(
  "https://github.com/mdsumner/aad-filelist/releases/download/file_db/raad_file_db.parquet"
)
raad_df |> dplyr::mutate(url = paste0("https://", file))
```

---

*Synced via GitHub Actions.
Source: [`mdsumner/aad-filelist`](https://github.com/mdsumner/aad-filelist)*
